Active Typechecking and Translation: A Safe Language-Internal Extension Mechanism

### 2013-10-12 esop14/att-esop14.tex
all of this talk about control was really unclear to me. i don't think it added much
without it, i think it's a whole lot better. actually, removing the
"without allowing so much control that the reliability of the system
suffers" part made the two sentences connect much better. besides, i
think that that subsentence is a bit hand wavy anyway, could piss
reviewers off more than just omitting it

A more appropriate and useful place for extensions like this is directly within libraries, alongside abstractions that do not require this level of control. To enable this, the language must allow for the introduction new primitive type families, like \lstinline{Prod}, operators, like \lstinline{Pair}, \lstinline{PrL} and \lstinline{PrR}, and associated typechecking and translation logic, without allowing so much control that the reliability of the system suffers. When encountering these new operators in expressions, the compiler must effectively  hand control over typechecking and translation to the appropriate user-defined logic. Because this mechanism is {language-internal}, all compilers must support it to satisfy the language specification.

>>>>

A more appropriate and useful place for extensions like this is directly within libraries. To enable this, the language must allow for the introduction new primitive type families, like \lstinline{Prod}, operators, like \lstinline{Pair}, \lstinline{PrL} and \lstinline{PrR}, and associated typechecking and translation logic. 
% without allowing so much control that the reliability of the system suffers. 
When encountering these new operators in expressions, the compiler must effectively hand control over typechecking and translation to the appropriate user-defined logic. Because this mechanism is {language-internal}, all compilers must support it to satisfy the language specification.
